---
title: "CSDA 1010 - Lab2 - Group 1"
output:
  html_document:
    df_print: paged
---

#### Importing libraries

```{r}
library(plyr); library(dplyr)
library(Hmisc)
library(corrplot)
library(caret)
library(pROC)
library(ROCR)
library(Metrics)
library(caTools)
library(rpart.plot)
library(boot)
```

#### Importing dataset

```{r}
data <- read.csv('online_shoppers_intention.csv', header = TRUE)
```

## Data understanding (Section 2 of the report)

### Dataset structure 

```{r}
str(data)
```

### Dataset summary

*examining basic descriptive statistics with "summary" function*
```{r}
summary(data)
```

### Checking for missing values

```{r}
colSums(is.na(data))
```

```{r}
colSums(data=="")
```
*dataset does not contain missing values*

### Feature visualizations (Only select visualizations were used/discussed in the report)

#### Revenue

*extracting and plotting the frequency of target feature "Revenue"*
```{r}
freq_Revenue=table(data$Revenue)
head(freq_Revenue)
```
```{r}
barplot(freq_Revenue)
```

*frequencies are converted to proportions, for better understanding and visualization*
```{r}
prop.table(freq_Revenue)
```
```{r}
barplot(prop.table(freq_Revenue))
```

#### Administrative

*plotting the histogram of "Administrative" feature; analogous plots will be used for other features in continuation, where appropriate*
```{r}
hist(data$Administrative, breaks=seq(0,27,1), labels=FALSE)
```

*visualizing proportions of "Adminstrative" feature against the target "Revenue" feature; analogous plots will be used for other features in continuation*
```{r}
freq_Administrative=xtabs(~data$Revenue+data$Administrative)
prop.table(freq_Administrative)
```
```{r}
barplot(prop.table(freq_Administrative),legend=rownames(freq_Administrative))
```

*plotting the mosaic plot of frequencies of "Administrative" vs "Revenue" feature, with shaded residual levels for cell/level combinations of the two features; analogous plots will be used for other features in continuation*  
```{r}
mosaicplot(freq_Administrative, border = "black",
           shade = TRUE)
```

#### Administrative_Duration

```{r}
hist(data$Administrative_Duration, breaks=seq(0,3400,100))
```

*breaking down the histogram of "Administrative_Duration" based on the "Revenue" feature; analogous plots will be used in continuation, as appropriate*
```{r}
data %>% 
  ggplot() +
  aes(x = Administrative_Duration) +
  geom_histogram(bins = 100, color='darkblue', fill='gray') +
  facet_grid(Revenue ~ .,
             scales = "free_y")
```

#### Informational

```{r}
hist(data$Informational, breaks=seq(0,24,1), labels=FALSE)
```

```{r}
freq_Informational=xtabs(~data$Revenue+data$Informational)
prop.table(freq_Informational)
```

```{r}
barplot(prop.table(freq_Informational),legend=rownames(freq_Informational))

```

```{r}
mosaicplot(freq_Informational, border = "black",
           shade = TRUE)
```

#### Informational_Duration

```{r}
hist(data$Informational_Duration, breaks=seq(0,2600,100))
```

```{r}
data %>% 
  ggplot() +
  aes(x = Informational_Duration) +
  geom_histogram(bins = 100, color='darkblue', fill='gray') +
  facet_grid(Revenue ~ .,
             scales = "free_y")
```

#### ProductRelated

```{r}
hist(data$ProductRelated,breaks=seq(0,750,50))
```

```{r}
data %>% 
  ggplot() +
  aes(x = ProductRelated) +
  geom_histogram(bins = 100, color='darkblue', fill='gray') +
  facet_grid(Revenue ~ .,
             scales = "free_y")
```

#### ProductRelated_Duration

```{r}
hist(data$ProductRelated_Duration, breaks=seq(0,64000,1000))
```

*breaking down the histogram of "ProductRelated_Duration" based on the "Revenue" feature to check how whether there is a difference of time spent on product pages in sessions that ended up with transaction*
```{r}
data %>% 
  ggplot() +
  aes(x = ProductRelated_Duration) +
  geom_histogram(bins = 100, color='darkblue', fill='gray') +
  facet_grid(Revenue ~ .,
             scales = "free_y")
```

*calculating the range of each bin on the x-axis"*
```{r}
(data %>%
  summarise(max_ProductRelated = max(data$ProductRelated_Duration, na.rm = TRUE)))/100
```

#### Bounce rates

```{r}
hist(data$BounceRates)
```

```{r}
data %>% 
  ggplot() +
  aes(x = BounceRates) +
  geom_histogram(bins = 100, color='darkblue', fill='gray') +
  facet_grid(Revenue ~ .,
             scales = "free_y")
```

#### Exit rates

```{r}
hist(data$ExitRates)
```

```{r}
data %>% 
  ggplot() +
  aes(x = BounceRates) +
  geom_histogram(bins = 100, color='darkblue', fill='gray') +
  facet_grid(Revenue ~ .,
             scales = "free_y")
```

#### Page values

```{r}
hist(data$PageValues)
```

```{r}
data %>% 
  ggplot() +
  aes(x = PageValues) +
  geom_histogram(bins = 100, color='darkblue', fill='gray') +
  facet_grid(Revenue ~ .,
             scales = "free_y")
```

#### Special day

```{r}
freq_SpecialDay=xtabs(~data$Revenue+data$SpecialDay)
prop.table(freq_SpecialDay)
```

```{r}
barplot(prop.table(freq_SpecialDay),legend=rownames(freq_SpecialDay))
```

```{r}
mosaicplot(freq_SpecialDay, border = "black",
           shade = TRUE)
```

#### Operating Systems

```{r}
freq_OperatingSystems=xtabs(~data$Revenue+data$OperatingSystems)
prop.table(freq_OperatingSystems)
```

```{r}
barplot(prop.table(freq_OperatingSystems),legend=rownames(freq_OperatingSystems))
```

```{r}
mosaicplot(freq_OperatingSystems, border = "black",
           shade = TRUE)
```

#### Browser

```{r}
freq_Browser=xtabs(~data$Revenue+data$Browser)
prop.table(freq_Browser)
```

```{r}
barplot(prop.table(freq_Browser),legend=rownames(freq_Browser))
```

```{r}
mosaicplot(freq_Browser, border = "black",
           shade = TRUE)
```

#### Region

```{r}
freq_Region=xtabs(~data$Revenue+data$Region)
prop.table(freq_Region)
```

```{r}
barplot(prop.table(freq_Region),legend=rownames(freq_Region))
```

```{r}
mosaicplot(freq_Region, border = "black",
           shade = TRUE)
```

#### Traffic Type

```{r}
freq_TrafficType=xtabs(~data$Revenue+data$TrafficType)
prop.table(freq_TrafficType)
```

```{r}
barplot(prop.table(freq_TrafficType),legend=rownames(freq_TrafficType))
```

```{r}
mosaicplot(freq_TrafficType, border = "black",
           shade = TRUE)
```

#### Visitor Type

```{r}
freq_VisitorType=xtabs(~data$Revenue+data$VisitorType)
prop.table(freq_VisitorType)
```


```{r}
barplot(prop.table(freq_VisitorType),legend=rownames(freq_VisitorType))
```

```{r}
mosaicplot(freq_VisitorType, border = "black",
           shade = TRUE)
```

#### Weekend

```{r}
freq_Weekend=xtabs(~data$Revenue+data$Weekend)
prop.table(freq_Weekend)
```

```{r}
barplot(prop.table(freq_Weekend),legend=rownames(freq_Weekend))
```

```{r}
mosaicplot(freq_Weekend, border = "black",
           shade = TRUE)
```

#### Month

```{r}
freq_Month=xtabs(~data$Revenue+data$Month)
prop.table(freq_Month)
```

```{r}
barplot(prop.table(freq_Month),legend=rownames(freq_Month))
```

```{r}
mosaicplot(freq_Month, border = "black",
           shade = TRUE)
```


## FEATURE ENGINEERING & MODELLING (Used in Sections 3 and 4 of the report)

### Transforming the dataset for Decision tree


```{r}
dataTree = data
```

*Changing categorical variables to ordered factors*
```{r}
dataTree$OperatingSystems <- factor(dataTree$OperatingSystems, order = TRUE, levels = c(6,3,7,1,5,2,4,8))
dataTree$Browser <- factor(dataTree$Browser, order = TRUE, levels = c(9,3,6,7,1,2,8,11,4,5,10,13,12))
dataTree$Region <- factor(dataTree$Region, order = TRUE, levels = c(8,6,3,4,7,1,5,2,9))
dataTree$TrafficType <- factor(dataTree$TrafficType, order = TRUE, levels = c(12,15,17,18,13,19,3,9,1,6,4,14,11,10,5,2,20,8,7,16))
dataTree$Month <- factor(dataTree$Month, order = TRUE, levels =c('Feb', 'Mar', 'May', 'June','Jul', 'Aug', 'Sep','Oct', 'Nov','Dec'))
dataTree$VisitorType <- factor(dataTree$VisitorType, order = TRUE, levels = c('Returning_Visitor', 'Other', 'New_Visitor'))
```


```{r}
str(dataTree)
```

### DECISION TREE MODEL 

*Creating train and test datasets with the ratio of 0.8*
```{r}
spl = sample.split(dataTree$Revenue, SplitRatio = 0.8)
train_tree = subset(dataTree, spl==TRUE)
test_tree = subset(dataTree, spl==FALSE)
```

*Building decision tree using "rpart"*
```{r}
dtree <- rpart(Revenue~., data = train_tree, method = 'class')
rpart.plot(dtree, extra = 'auto')
```

*Validate decision tree using complexity parameter (cp) and cross-validation error (xerror)*
```{r}
printcp(dtree)
plotcp(dtree)
```

*Confirm tree has been correctly pruned by manually selecting CP with minimum xerror*
```{r}
ptree<- prune(dtree, cp= dtree$cptable[which.min(dtree$cptable[,'xerror']),'CP'])
rpart.plot(ptree)
```

*Figure above shows the tree had been pruned by rpart function*

*Create prediction variable for test data*
```{r}
predTree <-predict(ptree, test_tree, type = 'class')

```

*Confusion Matrix*
```{r}
confmat_Tree <- table(test_tree$Revenue, predTree)
confmat_Tree
```

*Precision*
```{r}
precision_Tree <- confmat_Tree[2,2]/(confmat_Tree[2,2] + confmat_Tree[1,2])
precision_Tree
```

*Recall*
```{r}
recall_Tree <- confmat_Tree[2,2]/ (confmat_Tree[2,2] + confmat_Tree[2,1])
recall_Tree
```

*Accuracy*
```{r}
accuracy_Tree <- (confmat_Tree[2,2]+confmat_Tree[1,1])/ nrow(test_tree)
accuracy_Tree
```

*Balanced Accuracy*
```{r}
balancedacc_Tree <- (confmat_Tree[1,1]/(confmat_Tree[1,1] + confmat_Tree[1,2]) + confmat_Tree[2,2]/ (confmat_Tree[2,1] + confmat_Tree[2,2]))/2
balancedacc_Tree
```

### Transforming the dataset for Regressions

*Converting logical variables to binary dummy variables*
```{r}
dataReg <- data %>%
  mutate(Revenue_binary = ifelse(Revenue == "FALSE",0,1)) %>%
  mutate(Weekend_binary = ifelse(Weekend == "FALSE",0,1)) %>%
  select (-c(Revenue, Weekend))
```

*Converting factor to numeric variable*
```{r}
dataReg$Month <- as.numeric(dataReg$Month)
dataReg$VisitorType <- as.numeric(dataReg$VisitorType)
```

```{r}
str(dataReg)
```

*Creating correlation matrix*
```{r}
cor_result=rcorr(as.matrix(dataReg))
R2 <-  cor_result$r
```
```{r}
cor_result$r
```

```{r}
corrplot(R2, method="circle")
```

*Creating train and test sets for regression models using ratio of 0.7*
```{r}
spl2 = sample.split(dataReg$Revenue, SplitRatio = 0.7)
trainReg = subset(dataReg, spl2==TRUE)
testReg = subset(dataReg, spl2==FALSE)
```

### LINEAR REGRESSION MODEL

*First linear regression including all of the datasets variables*
```{r}
lm1 <-lm(ProductRelated_Duration ~., data = dataReg) 
summary(lm1)

```

*Creating a second linear regression that removed the statisticly insignificant results from the original regression and made the months a binary variable for each month*
```{r}
lm2 <- lm(ProductRelated_Duration ~ Administrative + Administrative_Duration + Informational_Duration + ProductRelated + BounceRates + ExitRates + SpecialDay + factor(Month) + VisitorType + Weekend_binary + Revenue_binary, data = dataReg)
summary(lm2)

par(mfrow=c(2,2))
plot (lm2)
```

*Creating a train model*
```{r}
trainlm2 <-lm(ProductRelated_Duration ~ Administrative + Administrative_Duration + Informational_Duration + ProductRelated + BounceRates + ExitRates + SpecialDay + factor(Month) + VisitorType + Weekend_binary + Revenue_binary, data = trainReg)
summary(trainlm2)
```

*Testing the model*
```{r}
testlm2 <-predict(trainlm2, testReg)
lm2rmse <- rmse(actual = testReg$ProductRelated_Duration,predicted = testlm2)
print(lm2rmse)
```

*Testing the new model against the first model using the same train and test data*
```{r}
trainlm1 <- lm(ProductRelated_Duration ~., data = trainReg)

testlm1 <-predict(trainlm1, testReg)
lm1rmse <- rmse(actual = testReg$ProductRelated_Duration,predicted = testlm1)
print(lm1rmse)
```

### LOGISTIC REGRESSION MODEL

*Building the model on training data*
```{r}
glm = glm(Revenue_binary ~ .,data=trainReg, family=binomial)
summary(glm)
```

*Prediction on the training set*
```{r}
predictglm = predict(glm, type="response")
summary(predictglm)
```

*Using tapply function to compute the average prediction for each of the outcomes*
```{r}
tapply(predictglm, trainReg$Revenue_binary, mean)
```

*Selecting a treshold value*
```{r}
ROCRpred = prediction(predictglm, trainReg$Revenue_binary)

```

*Calculating performace function*
```{r}
ROCRperf = performance(ROCRpred, "tpr", "fpr")

```

*Plotting ROC curve with treshold labels*
```{r}
ROCRperf = performance(ROCRpred, "tpr", "fpr")
plot(ROCRperf, colorize=TRUE)
plot(ROCRperf, colorize=TRUE, print.cutoffs.at=seq(0,1,by=0.1), text.adj=c(-0.2,1.7))

```

*Prediction on test set using the selected treshold*
```{r}
predictTest = predict(glm, type = "response", newdata = testReg)
```

```{r}
table(testReg$Revenue_binary,predictTest >= 0.3)
```

```{r}
confmat_Glm <- table(testReg$Revenue_binary,predictTest >= 0.3)
confmat_Glm

```

*Precision*
```{r}
precision_Glm <- confmat_Glm[2,2]/(confmat_Glm[2,2] + confmat_Glm[1,2])
precision_Glm
```

*Recall*
```{r}
recall_Glm <- confmat_Glm[2,2]/ (confmat_Glm[2,2] + confmat_Glm[2,1])
recall_Glm
```

*Accuracy*
```{r}
accuracy_Glm <- (confmat_Glm[2,2]+confmat_Glm[1,1])/nrow(testReg)
accuracy_Glm
```

*Balanced Accuracy*
```{r}
balancedacc_Glm <- (confmat_Glm[1,1]/(confmat_Glm[1,1] + confmat_Glm[1,2]) + confmat_Glm[2,2]/ (confmat_Glm[2,1] + confmat_Glm[2,2]))/2
balancedacc_Glm
```

*Root mean square error*
```{r}
testReg = testReg %>%
  mutate(predictions_quad = predict(glm, testReg))

sqrt(testReg %>%
  summarise(RMSE_glm = mean((Revenue_binary-predictions_quad)^2)))
```


